---
title: "03_Beta_diversity_analysis"
author: "Annegreet Veeken"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
  html_document:
    code_folding: hide
    toc: true
editor_options:
  markdown:
    wrap: 72
---

```{r setup}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
## Load packages
if(!require(tidyverse)) install.packages(tidyverse)
if(!require(geosphere)) install.packages(geosphere)
if(!require(easystats)) install.packages(easystats)
if(!require(rnaturalearth)) install.packages(rnaturalearth)
if(!require(rnaturalearthdata)) install.packages(rnaturalearthdata)
if(!require(lme4)) install.packages(lme4)
if(!require(spdep)) install.packages(spdep)
if(!require(sf)) install.packages(sf)
if(!require(broom)) install.packages(broom)
if(!require(MuMIn)) install.packages(MuMin)
if(!require(patchwork)) install.packages(patchwork)
if(!require(iNEXT.beta3D)) install.packages(iNEXT.beta3D)
if(!require(terra)) install.packages(terra)

## Load data
dir <- "~/Data/Globnut_offline/_GLOBNUT1.0/"

globnut <- readRDS("outputs/02_GlobNut.rds") %>% 
  ungroup() %>% 
  # dplyr::select(plot_ID, cell,cell_lon, cell_lat, habitat, lon, lat, sample_year,
  #               ndep, MAP, MAT, PET, NP, biomass, elev) %>% 
  # grouping for beta diversity is habitat by cell
  mutate(cell_hab = paste(cell, habitat, sep = "_"))
spec_raw <- read.csv(paste0(dir, "GlobNut1.0_species.csv")) 
spec <- spec_raw %>% 
  ungroup() %>% 
  # only vascular plants
  dplyr::filter(vascular_plant == 1) %>% 
  # convert subspecies to species
  mutate(species_new = 
           str_remove(species_new, pattern = " subsp.*| var\\.")) %>%
  group_by(plot_ID, species_new) %>% # aggregate by species
  summarise(cover = sum(cover, na.rm = TRUE)) %>% 
  dplyr::filter(cover != 0)  %>% 
  #  to 100 %
  group_by(plot_ID) %>% 
  mutate(covertot = sum(cover), 
         cover = round(cover/covertot*100))  %>% 
  left_join(globnut %>% dplyr::select(plot_ID, cell, habitat, cell_hab), by = "plot_ID") %>% 
  drop_na(cell, habitat, cell_hab) %>% 
  filter(cover > 0)

set.seed(123)
nobs <- 3
```

```{r functions}
result_figure <- function(mod) {
  p <- mod %>% 
   model_parameters() %>%   
  filter(Effects == "fixed") %>% 
  filter(!Parameter %in% c("(Intercept)", "distm")) %>% 
  mutate(Parameter = case_match(Parameter,
                "dif_np" ~ "Difference N/P",
                "ndep" ~ "N deposition",
                "mean_np" ~ "Mean N/P",
                "dif_bio" ~ "Difference biomass",
                "mean_bio" ~ "Mean biomass")) %>% 
  ggplot(aes(y = Parameter, x = Coefficient)) +
    geom_vline(aes(xintercept = 0)) +
    geom_errorbar(aes(xmin = CI_low, xmax = CI_high),  width = 0.25) +
  geom_point(shape = 21, size = 3, fill = "#a1d99b") +        
  scale_x_continuous("Standardized effect") +
  scale_y_discrete("") + 
  theme_bw()+
  theme(legend.position = "none")
p
}

result_table <- function(mod){
  tab <-  mod %>% 
  model_parameters() %>%   
  filter(Effects == "fixed") %>% 
  mutate(Parameter = case_match(Parameter,
                           "(Intercept)" ~ "Intercept",
                "dif_np" ~ "Difference N/P",
                "ndep" ~ "N deposition",
                "mean_np" ~ "Mean N/P",
                "dif_bio" ~ "Difference biomass",
                "distm" ~ "Geographical distance",
                "mean_bio" ~ "Mean biomass")) %>% 
  # Combine coefficient and SE
  mutate(`Estimate (SD)`= paste0(round(Coefficient, 3), " (", round(SE, 3), ")"),
      # Combine CI
      CI = paste0("[", round(CI_low, 3), ", ", round(CI_high, 3), "]"),
      `Significance level` = case_when(
        p < 0.001 ~ "***",
                    p < 0.01 ~ "**",
                    p < 0.05 ~ "*",
                    TRUE ~ ""),
      `p-value` = round(p, 4)) %>% 
  select(Parameter, `Estimate (SD)`, CI, `p-value`, `Significance level`) %>% 
  add_row(Parameter = c("Conditional R2", "Marginal R2"),
          `Estimate (SD)` = c(round(r2(mod)$R2_conditional,3), round(r2(mod)$R2_marginal,3)) %>% as.character()) %>% 
  knitr::kable()
  return(tab)
}
```

```{r ecoregions}

# # Extract ecoregions for globnut
# ecoregions <- terra::vect("~/Data/Ecoregions2017/Ecoregions2017.shp")
# # terra::crs(ecoregions)
# coords_df <- globnut %>%
#   dplyr::select(cell, cell_lon, cell_lat) %>%
#   distinct()
# coords <- terra::vect(coords_df, geom = c("cell_lon", "cell_lat"),  crs = "EPSG:4326")
# ecoreg_globnut <- cbind(coords_df[,c("cell", "cell_lon", "cell_lat")],terra::extract(ecoregions,  coords)) %>% 
#   select(cell, cell_lon, cell_lat, ECO_NAME)
# 
# # take the nearest for cells that didn't return a value
# missing_values <- ecoreg_globnut %>% 
#   filter(is.na(ECO_NAME))
# missing_coords_df <- missing_values %>% 
#   dplyr::select(cell, cell_lon, cell_lat)
# missing_coords <- terra::vect(missing_coords_df, geom = c("cell_lon", "cell_lat"),  crs = "EPSG:4326")
# nearest_ecoreg <- cbind(missing_coords_df[,c("cell", "cell_lon", "cell_lat")], terra::nearest(missing_coords, ecoregions))
# ids <- terra::values(ecoregions) %>% 
#   as.data.frame() %>% 
#   dplyr::select(OBJECTID, ECO_NAME) 
# missing_ecoreg <- nearest_ecoreg %>% 
#   left_join(ids, by = c("to_id" = "OBJECTID")) %>% 
#   select(cell, cell_lon, cell_lat, ECO_NAME)
# 
# # combine 
# ecoreg_globnut <- ecoreg_globnut %>% 
#   filter(!is.na(ECO_NAME)) %>% 
#   bind_rows(missing_ecoreg)
  
# saveRDS(ecoreg_globnut, "outputs/Ecoregions_emep_grid.rds")
ecoreg <- readRDS( "outputs/Ecoregions_emep_grid.rds")
```

```{r nobs}
# in cell
# check diff in sample year
# discard plot pairs sampled > 10 year  from eachother
plot_ids <- globnut %>% 
  group_by(cell) %>% 
  # all combinations of plots in a cell
  tidyr::expand(plot_x = plot_ID, plot_y = plot_ID) %>% 
  left_join(globnut %>% select(plot_ID, sample_year), by = c("plot_x" = "plot_ID")) %>% 
  left_join(globnut %>% select(plot_ID, sample_year), by = c("plot_y" = "plot_ID"), 
            suffix = c("_x","_y")) %>% 
  # amount of years between sampling
  mutate(diff_year = abs(sample_year_x - sample_year_y)) %>% 
  # discard plot pairs that have been sampled more than 10 years apart
  # select the  oldest plot
  mutate(oldest_plot = ifelse(sample_year_x < sample_year_y, plot_x, plot_y)) %>% 
  filter(!plot_x == plot_y) %>% # same plots
  filter(diff_year > 10) %>% 
  pull(oldest_plot) %>% 
  unique() 

## cells with enough data
cell_ids <- globnut %>% 
  # discard plots that are sampled too many years apart
  filter(!plot_ID %in% plot_ids) %>% 
  group_by(cell) %>% 
  summarise(n = n()) %>% 
  filter(n >= nobs) %>% 
  pull(cell)
```

There are `r length(cell_ids)` gridcells with at least `r nobs` plots

```{r map, warning=FALSE}
# Euroasia shape file
world <- ne_countries(scale = "medium", returnclass = "sf")
eurasia <- world[world$continent %in% c("Asia", "Europe"),]
coords_sf <- st_as_sf(unique(globnut[globnut$cell %in% cell_ids, c("cell_lon", "cell_lat")]), 
                      coords = c("cell_lon", "cell_lat"), crs = 4326)
  
p <- ggplot() +
  geom_sf(data = eurasia, fill = "grey90", col = "grey90") +
  geom_sf(data = coords_sf,  alpha = 0.25) +
  coord_sf(crs = "+proj=lcc +lat_1=45 +lat_2=55 +lat_0=50 +lon_0=35",
           xlim = c(-4000000, 4500000), ylim = c(-1000000, 4500000)) +
  theme_bw() 
p

```

## Method

Explanatory variables

-   N deposition per gridcell `ndep`

-   NP mean per gridcell `mean_np`

-   Mean difference in NP between plots in a gridcell `dif_np`

-   Biomass mean per gridcell `mean_bio`

-   Mean biomass difference between plots in a gridcell `dif_bio`

-   Mean MAT, MAP, PET in a gridcell `MAT, MAP, PET`

-   Mean pH `pH`

-   Mean distance in meters between plots `distm`

-   Mean difference in elevation between plots `elev`

-   Random effect of ecoregion `ECO_NAME`

Full model:
`beta ~ distm + elev + MAT + MAP + PET + pH + ndep + dif_np + mean_np + mean_bio + dif_bio + (1|ECO_NAME)`

Model selection on `distm + elev + MAT + MAP + PET + pH`, always
retaining `ndep + dif_np + mean_np + mean_bio + dif_bio + (1|ECO_NAME)`

Running Moran's test for checking residual spatial autocorrelation

```{r inext}
## Run iNext for rarefaction and extrapolation of beta diversity
# Convert to incidence format for inext
# incidence_data <- spec %>%
#   filter(!plot_ID %in% plot_ids) %>% 
#   filter(cell %in% cell_ids) %>%  # cells with at least nobs observations
#   filter(!cell == 538542) %>% # only 1 vasc species because high moss abundance, which iNext doesn't like
#   ungroup() %>% 
#   dplyr::select(cell, plot_ID, species_new, cover) %>%
#   mutate(species_new = str_replace(species_new, pattern = "[:space:]", replacement = "_")) %>% 
#   group_split(cell) %>%
#   purrr::map(., ~pivot_wider(.,names_from = "plot_ID",values_from = "cover", values_fill = 0) %>%
#                dplyr::select(.,-cell) %>% 
#                column_to_rownames("species_new") %>% 
#                as.matrix())

# beta_results <- iNEXTbeta3D(
#   data = incidence_data,
#   diversity = "TD",  # Taxonomic diversity
#   q = c(0, 1, 2),   # Multiple diversity orders
#   datatype = "abundance",
#   base = "coverage",
#   level = 0.9,
#   nboot = 200
# )
# saveRDS(beta_results, "outputs/inext_beta_output.rds")
beta_results <- readRDS("outputs/inext_beta_output.rds")
# Extract beta diversity estimates
beta0 <- beta_results %>% 
  map_df(., "beta") %>% 
  filter(Order.q == 0) %>% 
  mutate(cell = cell_ids[!cell_ids == 538542]) %>% 
  rename(beta_inext = Beta)
# Extract beta diversity estimates
beta1 <- beta_results %>% 
  map_df(., "beta") %>% 
  filter(Order.q == 1) %>% 
  mutate(cell = cell_ids[!cell_ids == 538542]) %>% 
  rename(beta_inext = Beta)
# Extract beta diversity estimates
beta2 <- beta_results %>% 
  map_df(., "beta") %>% 
  filter(Order.q == 2) %>% 
  mutate(cell = cell_ids[!cell_ids == 538542]) %>% 
  rename(beta_inext = Beta)
```

```{r env-var}
## Calculating explanatory variables
env_var <- globnut %>% 
 # same plots and cells as the incidence data
 filter(!plot_ID %in% plot_ids) %>% 
 filter(cell %in% cell_ids) %>% 
 group_by(cell) %>% 
 summarise(
   nplots = n_distinct(plot_ID),
   ## Confounders
   # Distance between plots
   distm = mean(distm(cbind(lon, lat), fun = distGeo)),
           # Elevation difference between plots
           elev = {diff_matrix <- abs(outer(elev, elev, "-"))
           mean(diff_matrix[lower.tri(diff_matrix)])
           },
           # Average climate
           MAT = mean(MAT, na.rm = TRUE),
           MAP = mean(MAP, na.rm = TRUE),
           PET = mean(PET, na.rm = TRUE),
           # Number of habitats in a cell
           nhab = n_distinct(habitat),
           # average pH
           pH = mean(pH),
           ## Predictors
           # n deposition
           ndep = mean(ndep),
           # Average biomass
           mean_bio = mean(biomass),
           # Average N/P
           mean_np = mean(NP),
           # Differences in NP
           dif_np = {diff_matrix <- abs(outer(NP, NP, "-"))
           mean(diff_matrix[lower.tri(diff_matrix)])
           },
           # Differences in biomass
           dif_bio =  {diff_matrix <- abs(outer(biomass, biomass, "-"))
             mean(diff_matrix[lower.tri(diff_matrix)])
           },
           # sample year
           dif_year = {diff_matrix <- abs(outer(sample_year, sample_year, "-"))
           mean(diff_matrix[lower.tri(diff_matrix)])
           }
           ) %>% 
 mutate(ndep = log(ndep*10),
        mean_bio = log(mean_bio),
        dif_bio = log(dif_bio),
        dif_np = log(dif_np),
        mean_np = log(mean_np),
        distm = log(distm +1),
        elev = log(elev+1)
        ) %>% 
 drop_na()
  
# hist(env_var$elev)
# hist(env_var$distm)
# hist(env_var$MAT)
# hist(env_var$MAP)
# hist(env_var$PET)
# hist(env_var$pH)
# hist(env_var$ndep)
# hist(env_var$mean_np)
# hist(env_var$mean_bio)
# hist(env_var$dif_bio)
# hist(env_var$dif_np)
# hist(env_var$nhab)
```

```{r mod-data}
mod_dat <- env_var %>%    
  # cell coordinates
  left_join(globnut %>% 
              dplyr::select(cell, cell_lon, cell_lat) %>% 
              distinct(), by = "cell") %>%
  # ecoregions
  left_join(ecoreg %>% dplyr::select(cell, ECO_NAME), by = "cell") %>% # Check missing cells ecoregions!!!
  # standardized beta div
  left_join(beta0 %>% dplyr::select(cell, beta_q0 = beta_inext), by = "cell") %>%
  left_join(beta1 %>% dplyr::select(cell, beta_q1 = beta_inext), by = "cell") %>% 
  left_join(beta2 %>% dplyr::select(cell, beta_q2 = beta_inext), by = "cell") %>% 
  drop_na() 
# hist(beta0$beta_inext, breaks = 20)
# hist(beta1$beta_inext, breaks = 20)
# hist(beta2$beta_inext, breaks = 20)
```

## q0

```{r}
# full model 
full_model_q0 <- lmer(log(beta_q0) ~ distm + elev + MAT + MAP + PET + pH  +
          ndep  + dif_np + mean_np + mean_bio + dif_bio + (1|ECO_NAME),
  data = mod_dat,
  na.action = "na.fail") 
# 
# check_model(full_model_q0) 
# summary(full_model_q0)

sb <- c("mean_bio", "ndep", "dif_bio", "mean_np", "dif_np")
dd <- dredge(full_model_q0, fixed =  sb)

selected_model_q0 <- lmer(get.models(dd,1)[[1]], data = mod_dat)
# check_model(selected_model_q0)
# summary(selected_model_q0)

# check residual spatial autocorrelation 
mod_resid <- residuals(selected_model_q0) 
knn_weights <- knearneigh(mod_dat[,c("cell_lon", "cell_lat")], k = 4)
knn_list <- knn2nb(knn_weights)
knn_weights_matrix <- nb2listw(knn_list, style = "W")

# Moran's I test 
moran_test <- moran.test(mod_resid, knn_weights_matrix)  

result_figure(selected_model_q0)
ggsave("figures/fig4-beta_diversity.png", dpi = 300, height = 4, width = 5)
result_table(selected_model_q0)

```

Moran's test p value:`r round(moran_test$p.value,4)` - no residual
autocorrelation

## q1

```{r}
# full model 
full_model_q1 <- lmer(log(beta_q1) ~ distm + elev + MAT + MAP + PET + pH  +
          ndep  + dif_np + mean_np + mean_bio + dif_bio + (1|ECO_NAME),
  data = mod_dat,
  na.action = "na.fail") 

# check_model(full_model_q1) 
# summary(full_model_q1)

sb <- c("mean_bio", "ndep", "dif_bio", "mean_np", "dif_np")
dd <- dredge(full_model_q1, fixed =  sb)

selected_model_q1 <- lmer(get.models(dd,1)[[1]], data = mod_dat)
# check_model(selected_model_q1)
# summary(selected_model_q1)

# check residual spatial autocorrelation 
mod_resid <- residuals(selected_model_q1) 
knn_weights <- knearneigh(mod_dat[,c("cell_lon", "cell_lat")], k = 4)
knn_list <- knn2nb(knn_weights)
knn_weights_matrix <- nb2listw(knn_list, style = "W")

# Moran's I test 
moran_test <- moran.test(mod_resid, knn_weights_matrix)  

result_figure(selected_model_q1)
result_table(selected_model_q1)

```

Moran's test p value:`r round(moran_test$p.value,4)` - still residual
autocorrelation

## q2

```{r}
# full model 
full_model_q2 <- lmer(log(beta_q2) ~ distm + elev + MAT + MAP + PET + pH  +
          ndep  + dif_np + mean_np + mean_bio + dif_bio + (1|ECO_NAME),
  data = mod_dat,
  na.action = "na.fail") 
# check_model(full_model_q2) 
# summary(full_model_q2)

sb <- c("mean_bio", "ndep", "dif_bio", "mean_np", "dif_np")
dd <- dredge(full_model_q2, fixed =  sb)

selected_model_q2 <- lmer(get.models(dd,1)[[1]], data = mod_dat)
# check_model(selected_model_q2)
# summary(selected_model_q2)

# check residual spatial autocorrelation 
mod_resid <- residuals(selected_model_q2) 
knn_weights <- knearneigh(mod_dat[,c("cell_lon", "cell_lat")], k = 4)
knn_list <- knn2nb(knn_weights)
knn_weights_matrix <- nb2listw(knn_list, style = "W")

# Moran's I test 
moran_test <- moran.test(mod_resid, knn_weights_matrix)  

result_figure(selected_model_q2)
result_table(selected_model_q2)

```

Moran's test p value:`r round(moran_test$p.value,4)` - still residual
autocorrelation
