---
title: "Biomass species compositon"
author: "Annegreet Veeken"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height=7,
                      fig.width=7,
                      message=FALSE, warning=FALSE)

## Load packages
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(lme4)) install.packages("lme4")
if(!require(performance)) install.packages("performance")
if(!require(parameters)) install.packages("parameters")
if(!require(ggeffects)) install.packages("ggeffects")
if(!require(multcompView)) install.packages("multcompView")
if(!require(MASS)) install.packages("MASS")
if(!require(knitr)) install.packages("knitr")
if(!require(DHARMa)) install.packages("DHARMa")

## Load data
globnut <- readRDS("outputs/02_GlobNut_subsampled.rds")
globnut <- globnut %>%
  mutate(lim = factor(lim, levels = c("No limitation by N, P, K", 
                                      "N-limitation",
                                      "P-limitation",
                                      "Co-limitation N-P"),
                      labels = c("No limitation by N, P, K", 
                                      "N limitation",
                                      "P limitation",
                                      "Colimitation N:P") ))
```

## Differences in biomass and species richness between nutrient limitation

We tested for differences in biomass between nutrient limitation using ANOVA and tukey HSD test.

```{r biomass ~ lim}
av <- aov(biomass ~ lim, data = globnut)
tuk <- TukeyHSD(av,"lim")
let <- multcompView::multcompLetters4(av,tuk)
let <- data_frame(lim = names(let$lim$Letters), sig_let = let$lim$Letters)
sig_lab <- globnut %>% 
  group_by(lim) %>% 
  summarise(mean = mean(biomass), sd = sd(biomass),
            n_obs = n()) %>% 
  left_join(let, by = "lim")

bio <- ggplot(globnut, aes(x = lim, y = biomass)) +
  geom_boxplot(fill = NA, width = 0.2) +
  geom_violin(alpha = 0.3) +
  geom_text(data = sig_lab, aes(x = lim, y = mean , label = sig_let)) +
  scale_x_discrete("") +
  scale_y_continuous("Biomass (g/m2)") +
  theme_bw()

```

```{r spec ric ~ lim }
av <- aov(spec_ric ~ lim, data = globnut)
tuk <- TukeyHSD(av,"lim")
let <- multcompView::multcompLetters4(av,tuk)
let <- data_frame(lim = names(let$lim$Letters), sig_let = let$lim$Letters)
sig_lab <- globnut %>% 
  group_by(lim) %>% 
  summarise(mean = mean(spec_ric), sd = sd(spec_ric),
            n_obs = n()) %>% 
  left_join(let, by = "lim")

sr <- ggplot(globnut, aes(x = lim, y = spec_ric)) +
  geom_boxplot(fill = NA, width = 0.2) +
  geom_violin(alpha = 0.3) +
  geom_text(data = sig_lab, aes(x = lim, y = mean , label = sig_let)) +
  scale_x_discrete("") +
  scale_y_continuous("Species richness") +
  theme_bw()
```

```{r grass cover ~ lim}
av <- aov(grass_cover ~ lim, data = globnut)
tuk <- TukeyHSD(av,"lim")
let <- multcompView::multcompLetters4(av,tuk)
let <- data_frame(lim = names(let$lim$Letters), sig_let = let$lim$Letters)
sig_lab <- globnut %>% 
  group_by(lim) %>% 
  summarise(mean = mean(grass_cover), sd = sd(grass_cover),
            n_obs = n()) %>% 
  left_join(let, by = "lim")

grass <- ggplot(globnut, aes(x = lim, y = grass_cover)) +
  geom_boxplot(fill = NA, width = 0.2) +
  geom_violin(alpha = 0.3) +
  geom_text(data = sig_lab, aes(x = lim, y = mean , label = sig_let)) +
  scale_x_discrete("") +
  scale_y_continuous("Grass cover (%)") +
  theme_bw()
```

```{r forb cover ~ lim}
av <- aov(forb_cover ~ lim, data = globnut)
tuk <- TukeyHSD(av,"lim")
let <- multcompView::multcompLetters4(av,tuk)
let <- data_frame(lim = names(let$lim$Letters), sig_let = let$lim$Letters)

sig_lab <- globnut %>% 
  group_by(lim) %>% 
  summarise(mean = mean(forb_cover), sd = sd(forb_cover),
            n_obs = n()) %>% 
  left_join(let, by = "lim")

forb <- ggplot(globnut, aes(x = lim, y = forb_cover)) +
  geom_boxplot(fill = NA, width = 0.2) +
  geom_violin(alpha = 0.3) +
  geom_text(data = sig_lab, aes(x = lim, y = mean, 
                                label = sig_let)) +
  scale_x_discrete("") +
  scale_y_continuous("Forb cover (%)") +
  theme_bw()

box_lim <- grid.arrange(bio, sr, grass, forb)
ggsave("figures/fig3-boxplots.png", box_lim,  width = 174, height = 174, units = "mm", dpi = 600)
```

## Species richness \~ Biomass relationship by nutrient limitation

-   Generalised linear model with linear and quadratic terms for biomass, plus their interaction with the nutrient limitations + plot size
-   negative binomial distribution

```{r echo=TRUE, fig.height=9}

spec_bio <- glm.nb(spec_ric ~ lim + biomass + I(biomass^2) +
                biomass*lim + I(biomass^2)*lim + plot_size, 
                data = globnut)

windows();check_model(spec_bio)

# test for overdispersion
testDispersion(spec_bio, plot = TRUE)
```

```{r fig.height=5}
ggpredict(spec_bio, terms = c("biomass [all]", "lim [all]")) %>% 
  plot(.) +
  ggtitle("") +
  xlab("Biomass (g/m2)") +
  ylab("Species richness") +
  theme(legend.title = element_blank())
ggpredict(spec_bio, terms = c("biomass", "lim")) %>% 
  plot(.,  show_residuals = TRUE) +
  scale_y_continuous(limits = c(0,200)) 
ggpredict(spec_bio, terms = "plot_size [all]") %>%
  plot(., show_residuals =  TRUE, show_residuals_line = TRUE)
ggplot(globnut, aes(x = biomass, y = spec_ric, 
                    col = plot_size)) +
  geom_point() +
  theme_bw()

glm_coef <- summary(spec_bio)$coefficients %>% 
  as.data.frame() %>% 
  rownames_to_column("parameter") %>% 
  mutate(across(where(is.numeric), ~round(., digits = 3))) %>% 
  mutate(sig = case_when(`Pr(>|z|)` < 0.001 ~ "***",
                         `Pr(>|z|)` < 0.01 ~ "**",
                         `Pr(>|z|)` < 0.05 ~ "*",
                         `Pr(>|z|)` < 0.1 ~ ".",
                         TRUE ~ ""))
kable(glm_coef)

```

R-square = `r r2_kullback(spec_bio) %>% round(digits = 2)`

```{r}
pred <- ggpredict(spec_bio, terms = c("biomass [all]", "lim [all]"))
sr <- pred %>% 
  group_by(group) %>% 
  summarise(max_sr = max(predicted) %>% round(),
            ci_low = nth(conf.low, which.max(predicted)) %>% round,
            ci_high = nth(conf.high, which.max(predicted)) %>% round,
            max_biomass = nth(x, which.max(predicted)) %>% round())
```

Following Palpurina et al. (2019) we tested if the species richness-biomass relationship is influenced by the type of nutrient limitation, now extending the biomass gradient from 500 to 1400 mg/m2. Irrespective of the type of nutrient limitation (except no limitation) we find the hump-backed shaped relationship between biomass and species richness (Grime). In line with Palpurina et al. (2019)?, the species richness peak is reached at lower biomass values for P limitation (`r sr$max_biomass[sr$group == "P limitation"]` mg/m2) than with co-limitation and N limitation (`r sr$max_biomass[sr$group == "Colimitation N:P"]` and `r sr$max_biomass[sr$group == "N limitation"]` mg/m2, respectively).\
Also in line with Palpurina et al. (2019), we see that the increase in species richness with biomass is steeper with co-limitation and P-limitation than with N-limitation. This leads to higher optimum in species richness in co-limited ecosystems than in the other types of nutrient limitation (`r sr$max_sr[sr$group == "Colimitation N:P"]` species, 95% confidence interval [`r sr$ci_low[sr$group == "Colimitation N:P"]` ,`r sr$ci_high[sr$group == "Colimitation N:P"]`]). The maximum species richness in P-limited ecosystems is lower than co-limited ecosystems (`r sr$max_sr[sr$group == "P limitation"]`, [`r sr$ci_low[sr$group == "P limitation"]` ,`r sr$ci_high[sr$group == "P limitation"]`]) and significantly higher than N-limited ecosystem (`r sr$max_sr[sr$group == "N limitation"]`, [`r sr$ci_low[sr$group == "N limitation"]` ,`r sr$ci_high[sr$group == "N limitation"]`]). We observe a decline in species richness with further increases in biomass. However, this decrease is stronger for co-limited and P-limited ecosystems than for N-limited ecosystems. This relative difference in decrease of species richness between co-limited and P-limited on one hand and N-limited ecosystems on the other ultimately leads to crossing the lines. Leading to the fact that highly productive ecosystem have higher species richness under N-limitation than under co-limitation or P-limitation
